\documentclass[10pt,aspectratio=169]{beamer}
\usetheme{default}
\usepackage[backend=biber,style=numeric]{biblatex}
\addbibresource{refs.bib}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[sfdefault]{roboto}
\renewcommand{\familydefault}{\sfdefault}

% Colors
\definecolor{primary}{RGB}{0,84,159}
\definecolor{secondary}{RGB}{142,186,229}
\setbeamercolor{palette primary}{bg=primary,fg=white}
\setbeamercolor{palette secondary}{bg=secondary,fg=white}
\setbeamercolor{progress bar}{fg=secondary}

% Presentation info
\title[Reevaluation of DialogueGCN]{\Large Reevaluation of DialogueGCN for Emotion Recognition in Conversation\\Faithful Reproduction, Adaptations and Practical Limitations}
\author[Hamady GACKOU]{Hamady GACKOU\\\small Supervisor: Severine AFFELT}
\institute{Universite Paris Cite \\ Master 1 AMSD}
\date{\today}

\begin{document}

% Title
\begin{frame}[plain]
  \titlepage
\end{frame}

% Outline
\begin{frame}{Outline}
  \tableofcontents
\end{frame}

% 1. Introduction
\section{Introduction}
\begin{frame}{Context and Challenges}
  \begin{itemize}
    \item \textbf{ERC}: understanding emotional dynamics in dialogues.
    \item Challenges: multi-speaker context, temporal evolution, linguistic ambiguity.
    \item Applications: empathetic assistants, human--machine interfaces, mental health.
  \end{itemize}
  \vfill
  \begin{block}{Objectives}
    \begin{itemize}
      \item Reproduce \emph{DialogueGCN} (Ghosal \emph{et al.}, EMNLP--IJCNLP 2019).
      \item Evaluate on IEMOCAP, MELD, DailyDialog.
      \item Explore hyperparameters and CPU vs GPU limitations.
    \end{itemize}
  \end{block}
\end{frame}

% 2. Related Work
\section{Related Work}
\begin{frame}{Sequential vs Graph-Based Approaches}
  \begin{columns}
    \column{0.48\textwidth}
    \begin{block}{RNN-based}
      \begin{itemize}
        \item LSTM/GRU: local context, loss of long-range information.
        \item DialogRNN: speaker-wise tracking \cite{majumder2019dialoguernn}.
      \end{itemize}
    \end{block}
    \column{0.48\textwidth}
    \begin{block}{Graph-based}
      \begin{itemize}
        \item DialogueGCN: multi-relational RGCN, modeling inter-/intra-speaker interactions \cite{ghosal2019dialoguegcn}.
        \item Variants: MHA-GCN (multimodal), RGCN++ (enhanced attention).
      \end{itemize}
    \end{block}
  \end{columns}
\end{frame}

% 3. Methodology
\section{Methodology}
\begin{frame}{Faithful Implementation}
  \begin{itemize}
    \item Official GitHub code (commit \texttt{6128ca2}).
    \item Architecture:
      \begin{itemize}
        \item 2-layer multi-relational RGCN.
        \item Context windows $(p,f)$ around each utterance.
        \item Active-listener mechanism, \texttt{general} attention.
      \end{itemize}
    \item Adaptations:
      \begin{itemize}
        \item Migration PyTorch~1.0 to 2.7.0, PyG~1.3 to 2.6.1.
        \item Modernized DataLoader, TensorBoard logging.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Hyperparameters and Environment}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{block}{Hyperparameters}
      \small
      \begin{tabular}{ll}
        LR & 1e-4 (L2=1e-5) \\
        Dropout (rec/std) & 0.1 / 0.5 \\
        Batch size & 32 \\
        Epochs & 60 \\
        Windows & (10,10) \\
        Attention & general \\
      \end{tabular}
    \end{block}
    \column{0.5\textwidth}
    \begin{block}{Environment}
      \begin{itemize}
        \item CPU: tests on 8\,GB RAM server.
        \item Libraries: PyTorch~2.7.0, PyG~2.6.1, Pandas~2.2.3.
      \end{itemize}
    \end{block}
  \end{columns}
\end{frame}

% 4. Experimental Protocol
\section{Experimental Protocol}
\begin{frame}{Datasets and Preprocessing}
  \begin{block}{Corpora}
    \begin{itemize}
      \item \textbf{IEMOCAP}: 151 dialogues, 11\,000 utterances, 6 classes.
      \item \textbf{MELD}: 1\,433 dialogues, 13\,708 utterances, 7 classes, multi-speaker.
      \item \textbf{DailyDialog}: 13\,118 dialogues, 102\,979 utterances, 7 classes.
    \end{itemize}
  \end{block}
  \vspace{0.5em}
  \begin{block}{Preprocessing}
    Unicode cleaning, label harmonization (excited$\to$happy), WordPiece (max 250 tokens), Pickle serialization.
  \end{block}
\end{frame}

% 5. IEMOCAP Results
\section{IEMOCAP Results}
\begin{frame}{Performance on IEMOCAP}
  \begin{table}
    \centering
    \begin{tabular}{lcc}
      \toprule
      Configuration & Original (F1) & Reproduction (F1) \\
      \midrule
      DialogueGCN & 64.18\% & 63.9\% \\
      + Class weights & -- & 61.54\% \\
      Dropout 0.3 vs 0.7 & -- & 58.38\% / 54.17\% \\
      LR 1e-4 vs 1e-3 & -- & 54.97\% / 60.43\% \\
      Context (5,5)/(10,10)/(15,15) & -- & 56.08\% / 59.11\% / 58.19\% \\
      \bottomrule
    \end{tabular}
    \caption{Comparison of configurations on IEMOCAP}
  \end{table}
  \vfill
  \begin{itemize}
    \item Best: LR=1e-3, dropout=0.3, window=(10,10).
    \item Active listener: +2 pts F1 on long dialogues.
  \end{itemize}
\end{frame}

% 6. MELD & DailyDialog
\section{MELD \& DailyDialog}
\begin{frame}{MELD vs DailyDialog}
  \begin{columns}
    \column{0.48\textwidth}
    \begin{block}{MELD}
      \begin{itemize}
        \item Original: 58.10\% F1.
        \item Reproduction: plateau at 48.12\% by epoch 1.
        \item CPU training: 240--280\,s/epoch.
      \end{itemize}
    \end{block}
    \column{0.48\textwidth}
    \begin{block}{DailyDialog}
      \begin{itemize}
        \item No paper baseline.
        \item Reproduction: 28.21\% $\to$ 82.28\% (step 340).
        \item Rapid convergence, stable dataset.
      \end{itemize}
    \end{block}
  \end{columns}
\end{frame}

% 7. Additional Studies
\section{Additional Studies}
\begin{frame}{Ablation \& Variations}
  \begin{itemize}
    \item \textbf{Attention}: dot-product $>$ general $>$ concat.
    \item \textbf{Batch size}: 16 $\to$ 57.56\%, 64 $\to$ 56.21\%.
    \item \textbf{Class weighting}: +4 pts F1 (61.54\%).
    \item \textbf{Context windows}: (10,10) optimal.
  \end{itemize}
  \vfill
  \begin{block}{CPU Limitation}
    Training on EmoryNLP impossible (707 nodes, 2105 edges).
  \end{block}
\end{frame}

% 8. Conclusion & Perspectives
\section{Conclusion and Perspectives}
\begin{frame}{Conclusion \& Perspectives}
  \begin{block}{Summary}
    \begin{itemize}
      \item Faithful reproduction validated, gaps <1\% on IEMOCAP.
      \item Hyperparameters and extensions are beneficial.
      \item CPU vs GPU limitations highlighted.
    \end{itemize}
  \end{block}
  \vspace{0.5em}
  \begin{block}{Future Work}
    \begin{itemize}
      \item Dynamic GCNs, edge filtering for memory efficiency.
      \item Multimodal fusion (audio, visual).
      \item Explainability (GNNExplainer) for transparency.
    \end{itemize}
  \end{block}
  \vfill
  \begin{center}
    \Large\color{primary} Thank you for your attention!
  \end{center}
\end{frame}

% References
\begin{frame}[allowframebreaks]{References}
    \printbibliography[heading=none]
\end{frame}

\end{document}
